Covering Your Basis: Model Agnostic Probes of the Binary Black Hole Population

### 2022-09-26 src/tex/appendicies.tex
Re-work intro, add citations etc

\appendix
\section{Hierarchical Bayesian Inference} \label{sec:hierarchical_inference}

@@

\noindent
Where up above, we denote the co-moving volume element as $dV_c$ \citep{hogg_cosmo}, and $T_\mathrm{obs}$ as the observing time period that produced the 
catalog with the related factor of $1+z$ converting this detector-frame time to source-frame. We assume a LambdaCDM cosmology using 
the cosmological parameters from \citet{Planck2015}. We model the merger rate evolving with redshift following a power law distribution: 
$p(z|\lambda) \propto \frac{dV_c}{dz}\frac{1}{1+z}(1+z)^\lambda$. When integrating equation \ref{number_density} across all $\theta$
and out to some maximum redshift, $z_\mathrm{max}$, we get the total number of CBCs in the universe out to that redshift. We follow previous notations, \
letting $\{d_i\}$ represent the set of data from $N_\mathrm{obs}$ CBCs observed with gravitational waves. The merger rate is then described as an inhomogenous 
Poisson process and after imposing the usual log-uniform prior on the merger rate, we marginalize out merger rate and arrive at the posterior
distribution of our hyper-parameters, $\Lambda$ \citep{Mandel_2019, Vitale_2021}.

@@

\noindent
where, we replaced the integrals over each event's likelihood with ensemble averages over $K_i$ posterior samples. Above, $j$
indexes the $K_i$ posterior samples from each event and $\pi(\theta, z)$ is the default prior used by parameter estimations that 
produced the posterior samples for each event. In the analyses of GWTC-2 the default prior used is uniform in detector frame masses, 
component spins and Euclidean volume. The corresponding prior evaluated in source frame primary mass, mass ratio, component spins 
and redshift is:

@@

\section{Penalized Splines and Smoothing}\label{sec:psplines}

Spline functions have been shown to be sensitivite to the chosen number of knots, and their locations or spacing. 
Adding more knots knots increases the a-priori variance in the spline function, while the space between knots can limit the 
resolution of features in the data the spline is capable of resolving. To ensure your spline based model is flexible enough 
one would want to add as many knots as densely as possible, but this comes with the inherit larger variance as mentioned above. 
To overcome this we resort to the frequentist penalized spline (P-Spline) in which one applies a penalty or regularization term 
to the likelihood based on the difference of adjacent knot coeficients. The linear combination of spline basis components is 
is flat when the basis coeficients are equal (see Figure \ref{fig:spline_basis}). By penalizing the likelihood as the differences between adjacent
knot coeficients get larger, one gets a smoothing of the spline function inferences. Since we are using hierarchical Bayesian inferece as our 
statistical framework, we formulate this penalized likelihood with its Bayesian analog. The difference penalty is equivalent to applying a stochastic 
differnece prior among the knots, as described in \bruce{CITE THIS PAPER}. This is also sometimes referred to as a Gaussian random walk prior, since
each coeficient is normally distributed about its neighboring coeficients. We then use as our smoothing prior, for a spline basis with $n$ degree's of freedom
, and a difference order of $r$ in our smoothing penalty (see \bruce{cite P-splines paper}) is defined as:

\begin{eqnarray}
\tau_\lambda \sim \Gamma(a, b) \\
p(\bm{c} | \tau_\lambda) \propto \exp \big[ -\frac{1}{2} \tau_\lambda \bm{c}^{\mathrm{T}} \bm{D_r}^{\mathrm{T}} \bm{D_r} \bm{c}  \big] e^{-n\bm{c}}
\end{eqnarray}

\noindent Above $\bm{D_r}$ is the order-$r$ difference matrix, of shape $(n-r \times n)$ and $\Gamma(a,b)$ the Gamma distribution with shape parameters, $a$ and $b$. 
This smoothing prior removes the strong dependence on number and location of knots that arises with using splines. By adding in a very large number 
of knots such that your domain is densly populated with basis coeficients, this allows the freedom for the model to find the smoothing 
scale that the data prefers. One downside of this prior specification as noted by\ bruce{CITE BAYESIAN PSPLINES}, is a slight dependence on choice 
of the b hyper-parameter for Gamma distribution prior on $\tau_\lambda$. We reduce this affect by following the method introduced in \bruce{CITE BAYESIAN PSPLINES} 
to alleviate this. This is done by modeling $\tau_\lambda$ with a mixture model on gamma distributions with varying b hyper-parameters. 
We now replace the prior on $\tau_\lambda$ with:

Since the smoothing prior is really wanting to pull the coeficients closer to each other to smooth the distribution, 
we must ensure that the spline function is in fact flat or smooth with all equal coeficients. Often times one would stack knots at the same values 
at hard parameter boundaries but this removes the feature that with equal weights the spline function is flat. We follow the distinction from \bruce{CITE psplines} 
that the difference penalty or prior is only valid with ``proper'' spline bases. A proper basis is where all of the knots are evenly spaced, see Figure \ref{fig:spline_basis}.

@@

\section{Validation Studies} \label{sec:validation}

>>>>

\appendix
\section{Penalized Splines and Smoothing}\label{sec:psplines}

Spline functions have been shown to be sensitivite to the chosen number of knots, and their locations or spacing. 
Adding more knots knots increases the a-priori variance in the spline function, while the space between knots can limit the 
resolution of features in the data the spline is capable of resolving. To ensure your spline based model is flexible enough 
one would want to add as many knots as densely as possible, but this comes with the inherit larger variance as mentioned above. 
To overcome this we resort to the frequentist penalized spline (P-Spline) in which one applies a penalty or regularization term 
to the likelihood based on the difference of adjacent knot coeficients. The linear combination of spline basis components is 
is flat when the basis coeficients are equal (see Figure \ref{fig:spline_basis}). By penalizing the likelihood as the differences between adjacent
knot coeficients get larger, one gets a smoothing of the spline function inferences. Since we are using hierarchical Bayesian inferece as our 
statistical framework, we formulate this penalized likelihood with its Bayesian analog. The difference penalty is equivalent to applying a stochastic 
differnece prior among the knots, as described in \bruce{CITE THIS PAPER}. This is also sometimes referred to as a Gaussian random walk prior, since
each coeficient is normally distributed about its neighboring coeficients. For a spline basis with $n$ degree's of freedom, 
and a difference penalty of order of $r$ (see \bruce{cite P-splines paper}), our prior is defined as:

\begin{eqnarray}
\tau_\lambda \sim \Gamma(a, b) \\
p(\bm{c} | \tau_\lambda) \propto \exp \big[ -\frac{1}{2} \tau_\lambda \bm{c}^{\mathrm{T}} \bm{D}_{r}^{\mathrm{T}} \bm{D}_r \bm{c}  \big] e^{-n\bm{c}}
\end{eqnarray}

\noindent Above $\bm{D}_r$ is the order-$r$ difference matrix, of shape $(n-r \times n)$ and $\Gamma(a,b)$ the Gamma distribution with shape parameters, $a$ and $b$. 
This smoothing prior removes the strong dependence on number and location of knots that arises with using splines. By adding in a very large number 
of knots such that your domain is densly populated with basis coeficients, this allows the freedom for the model to find the smoothing 
scale that the data prefers. One downside of this prior specification as noted by\ bruce{CITE BAYESIAN PSPLINES}, is a slight dependence on choice 
of the b hyper-parameter for Gamma distribution prior on $\tau_\lambda$. We reduce this affect by following the method introduced in \bruce{CITE BAYESIAN PSPLINES} 
to alleviate this. This is done by modeling $\tau_\lambda$ with a mixture model on gamma distributions with varying b hyper-parameters. 
We now replace the prior on $\tau_\lambda$ with:

This prior is imparting a natural attraction of the coeficients closer to each other in order to smooth the spline function, so one 
must ensure that the spline function is in fact flat given all equal coeficients. Since one needs to define n+k knots to construct an order-k basis with 
n degrees of freedom, people often will place knots on top of each other at hard parameter boundaries \bruce{cite msplines / o-splines}. 
This may seemed motivated, especially if there is no reality in the parameter beyond the boundary, but this violates the above condition. 
We follow the distinction from \bruce{CITE psplines} that the difference penalty or smooothing prior is only valid with ``proper'' spline bases. 
A proper basis is where all n+k knots are evenly spaced, see Figure \ref{fig:spline_basis}, as opposed to stacking them at the bounds.

\section{Hierarchical Bayesian Inference} \label{sec:hierarchical_inference}

@@

\noindent
Where up above, we denote the co-moving volume element as $dV_c$ \citep{hogg_cosmo}, and $T_\mathrm{obs}$ as the observing time period that produced the 
catalog with the related factor of $1+z$ converting this detector-frame time to source-frame. We assume a LambdaCDM cosmology using 
the cosmological parameters from \citet{Planck2015}. We model the merger rate evolving with redshift following a power law distribution: 
$p(z|\lambda) \propto \frac{dV_c}{dz}\frac{1}{1+z}(1+z)^\lambda$ \cite{Fishbach_2018redshift}. When integrating equation \ref{number_density} across all $\theta$
and out to some maximum redshift, $z_\mathrm{max}$, we get the total number of CBCs in the universe out to that redshift. We follow previous notations, \
letting $\{d_i\}$ represent the set of data from $N_\mathrm{obs}$ CBCs observed with gravitational waves. The merger rate is then described as an inhomogenous 
Poisson process and after imposing the usual log-uniform prior on the merger rate, we marginalize out merger rate and arrive at the posterior
distribution of our hyper-parameters, $\Lambda$ \citep{Mandel_2019, Vitale_2021}.

@@

\noindent
where, we replaced the integrals over each event's likelihood with ensemble averages over $K_i$ posterior samples. Above, $j$
indexes the $K_i$ posterior samples from each event and $\pi(\theta, z)$ is the default prior used by parameter estimations that 
produced the posterior samples for each event. In the analyses of GWTC-3, either the default prior used was uniform in detector frame masses, 
component spins and Euclidean volume or the posterior samples were re-weighted to such a prior before using them in our analysis. 
The corresponding prior evaluated in the parameters we hierarchicaly model, i.e. source frame primary mass, mass ratio, component spins and redshift is:

@@

@@

\bruce{Need to fill this section in and put tables with specific hyper-parameter/prior settings made}

%\section{Validation Studies} \label{sec:validation}

### 2022-09-26 src/tex/intro.tex
Re-work intro, add citations etc

Observations of gravitational waves from compact binary coalescences (CBCs) is becoming a regular occurrence, 
recently producing catalogs of 90 such detections \bruce{cite GWTCs}. This growing catalog size is beginning to enable access to a more complete 
understanding of the population properties CBCs, which encodes the history of the formation and evolution of these extreme compact objects. 
In the coming fourth observing run of the LIGO-Virgo-KAGRA collaboration \bruce{cite detectors}, the catalog size is expected to increase by up to a factor of 5. 
Looking at population properties of compact objects lets us probe the complex physics common to all of our observations that governs the 
formation and evolution of merging compact objects. The history of these dense objects is imprinted on the properties of the population 
of CBCs, allowing us to learn about these not well understood ideas through careful evaluation of the enture distribution of observed GWs. 
The mass distribution of binary black holes has been used to measure cosmological parameters, constrain modified gravitational wave propagation, 
constrain a non-constant Planck mass, look for evidence of ultralight bosons through superradiance, put constraints on the uncertain nuclear reaction 
rates important for stellar evolution, look for primordial black holes, to constrain the neutron star equation of state, 
and to understand the fromation channels/environments of compact binary coalescences (CBCs). Through a better understanding of the mass, spin, 
and redshift distribution of compact binaries that will come with the increased catalog size, one can probe these dufferent physical phenoma with 
greater fidelity.

With the latest LIGO-Virgo-KAGRA (LVK) catalog of CBCs (GWTC-3), new understanding of the population has come into light. One of the main findings was that 
the mass distribution exhibits substructure with peaks rather than being perfectly described by a smooth power law distribution. The evidence for this was 
found with mutliple different flexible or non-parametric models that aim to be data-oriented with little bias on the function form. While these models are 
useful for looking for unexpected features in the data, or finding where our simpler models fail, they can be harder to interpret and are not based on specific
astrophysical motivation. These types of models will become more important as the size of our catalogs increase and we enter a data-dominated regime rather than
prior/model dominated.

The spin distribution of binary black holes (BBHs) has also been of controversial interest in the literature recently, with different parameteric approaches finding 
rather different conclusions on the distribution of spins of BBHs which has significant impact on the inferences about relative contribution of dynamically formed
binaries versus ones formed in the field. This is another use case for flexible data-driven models to alleivate the discrepencies among parametric approaches. 
In particular there has been little work on flexible model development for the spin distribution of BBHs, which makes sense given the relatively smaller catalog 
sizes and the large uncertainty in individual black hole spin measurements with GWs.

In this paper we highlight how the use of basis-splines can be a useful model-agnostic modeling approach to the astrophysical distributions of compact 
binaries. We show that with a particular choice of basis, one can efficiently model the spin, mass and redshift distributions of compact binaries in GWTC-3 with
hierarchical Bayesian inference. We discuss our results in the context of current literature on compact object populations and illustrate how our results are similar
to those produced with the more simpler and astrophysically motivated parametric models. This data-drien approach will be capable of finding more subtle features, that 
may not be expected, should they appear in future catalogs. Additionally this method can be extended to higher-dimensional basis-splines in order to model 
joint population disributions over multiple surce parameters, which will be important to confidently disentangle different competing formation channels 
should they exist. In Section II we go over the background of basis splines and model construction. In Section III we present results using our basis spline model 
on the mass and spin distributions of BBHs in GWTC-3, followed by discussing the results and astrophysical implications in Section IV and concluding in Section V.

>>>>

Observations of gravitational waves (GWs) from compact binary coalescences (CBCs) is becoming a regular occurrence, 
recently producing catalogs of 90 such detection \cite{GWTC1,gwtc2,GWTC3}. With this growing catalog size, we are steadily approaching a more complete  
understanding of the population properties CBCs, just 7 years after the first observation \cite{o1o2_pop,o3a_pop,o3b_astro_dist}. 
In the coming fourth observing run of the LIGO-Virgo-KAGRA collaboration detectors \cite{aLIGO,aVIRGO, KAGRA}, with the increased sensitivities, 
there could be a detection as often as every day. This large influx of incoming data will drastically increase what we can learn by combining all observations
in order to studying the compact object population. The history of these dense objects is imprinted on the properties of the population 
of CBCs, allowing us to learn about their formation and evolution by the last few seconds of gravitation radiation emitted. The population distibutions 
of merging compact objects inferred from gravitational-wave observations has been shown to be parituclar useful in order to, measure cosmological parameters \citep{Farr_2019HUB}, 
constrain modified gravitational wave propagation, constrain a running Planck mass, look for evidence of ultralight bosons through superradiance, 
constrain stellar nuclear reaction rates \cite{Farmer_2020}, look for primordial black holes \cite{Ng_2021}, and constrain the neutron star equation of state \bruce{Cite all points}. 
Through a better understanding of the mass, spin, and redshift distributions of compact binaries that will come with the increased catalog size, 
one can probe a wide range of different physical phenoma with even greater fidelity.

Studies using flexible methods to analyze the binary black hole (BBH) mass distribution with GWTC-2 \citep{Tiwari_2021_b,Edelman_2022ApJ} found signs of structure 
beyond the reported $m_1\sim35\,\msun$ feature by \citet{Talbot_2019, o3a_pop}. The has evidence for this structure, such as the peak at $m_1\sim10\,\msun$, has only increased
with GWTC-3 \cite{o3b_astro_dist,Tiwari_2022ApJ}. This substructure was first confidently found with mutliple different flexible or non-parametric models that aim 
to be data-oriented with little bias on the functional form. Conversely the other canonical approach is to use simple parametric models that aim to describe the data
in the simplest way, building in astrophyscially motivated priors, followed by sequentially adding complexities only as the data demands it. While in the current environment 
this simple approach is clearly more powerful, as we go forward into the data-rich catalog era flexible models can be a useful complement.
Non-parametric or semi-parametric methods are particularly useful at looking for unexpected features in the data, or finding where 
our simpler parametric models may be missing features in the data. Development and robust teting of these types of methods in the short term will provide
knock dowm effects later down the line as the size of our catalogs increase and we enter a data-dominated regime rather than prior/model dominated.

The spin distributions of binary black holes (BBHs) have particular interest since the two main formation scenarios of BBHs produce 
distinctly different spin distrubitons \bruce{cite}. Isolated or field formation scenarios predict spins that are preferentially, although some small misalignment can occur 
depending on the nature of the supernovae kicks as each star collapses to a compact object \bruce{cite}. Dynamical formation scenarios occur in dense environments where 
many-body interactions between a large number compact objects occur before producing a binary that will merger within a Hubble time. Thus the dynamical channel produces 
an isotropic spin orientation distribution \bruce{cite}. It has also been of controversial interest in the literature recently, with different parameteric approaches finding 
different conclusions on the distribution of spins of BBHs. Recent studies have disagreed on the possible existence of a significant zero-spinning subpopulation, along with the 
presence of significant mis-alignment ($\cos{\theta_i} < 0.0$) \cite{o3b_astro_dist,RouletGWTC2Pop,BuildBetterSpinModels,GWTC3MonashSpin,Callister_NoEvidence}. A recent study has shown the spin orientation distribution infernce is senitive to modeling choices and 
may not peak at perfectly aligned spins, as is commonly assumed \cite{spinitasyoulike}. 
Other approaches to building better models for BBH spins again, as with the mass models, focus on iterating through different simple parameteric descriptions with 
increasing complexities \cite{BuildBetterSpinModels,Callister_NoEvidence,spinitasyoulike}. The spin distributions offer a nice use case for flexible data-driven models, both 
to help alleivate the current disputes in the literature and because there has been little work on flexible model development for the spin distribution of BBHs thus far.

In this paper we highlight how the use of basis-splines can be a useful non-parametric modeling approach to the astrophysical distributions of compact 
binaries. We show that with a particular choice of basis, one can efficiently model the mass and spin distributions of compact binaries in GWTC-3 with
hierarchical Bayesian inference. We discuss our results in the context of current literature on compact object populations and illustrate how our results are similar
to those produced with the more simpler and astrophysically motivated parametric models. This data-drien approach will be capable of finding more subtle features, that 
may not be expected, should they appear in future catalogs. In Section \ref{sec:basis_splines} we go over the background of basis splines, in Section \ref{sec:results} 
we present results using our basis spline model on the mass and spin distributions of BBHs in GWTC-3, followed by discussing results in their astrophysical context and concluding 
in Section \ref{sec:conclusion}.

### 2022-09-26 src/tex/methods.tex
Re-work intro, add citations etc

A common non-parametric method in statistical applications is the use of basis splines. A spline function of order $k$, 
is a piece-wise polynomial of order $k$ polynomials stitched together from defined ``knot'' locations across the domain. 
They provide a useful and cheap way to interpolate generically smooth functions from a finite sampling of ``knot'' heights. 
Basis splines of order $k$ are a set of order $k$ polynomials that form a complete basis for any spline function of order $k$. 
Given an array of knot locations, $\mathbf{t}$ or knot vector, there exists a single unique linear combination of B-Splines for 
every possible spline function with knots, $\mathbf{t}$. One can construct the B-Splines basis components for a given knot vector 
using the Cox-de Boor recursion formula. Given knots $t_0$, $t_1$,...,$t_{i+k}$ we start with the base case as:

@@

@@

\noindent allowing one to build a basis of $i$ components that spans the vector space of order-$k$ spline functions interpolated 
from this knot vector. This is known as the ``B-Spline'' basis.

A similar basis, the M-Spline basis, has desirable properties when looking to model a probability density function. Namely, 
$M_i \geq 0$ within $(t_i, t_i+k)$, zero elsewhere, and normalized to unity, $\int M_i(x)dx = 1$. The M-Spline basis only differs 
from B-Splines by a normalization factor, namely:

\noindent We use this definition of the M-Spline basis for the models presented and applied in the later sections of this paper.

\section{Model Construction}

\bruce{placeholder paragraph here to introduce our model constructions.}
We use hierarchical Bayesian inference to simultaneosly infer the mass, spin and redshift distributions of binary black holes (BBHs) using 
the 69 confident BBH detections reported in the recent LVK catalog of observations, GWTC-3. In particular we choose to parameterize the two
masses with the primary (more massive component) mass and the mass ratio defined as $q=\frac{m_2}{m_1}$. We choose to model 4 of the 6 total 
spin degrees of freedom of a binary merger with each component spin magnitude, and the cosine of the tilt angle which is defined as the angle 
between each spin vector and the binaries orbital angular mommentum vector. We additionally will fit a population model on the redshift distribution
of BBHs. We assume a $\Lambda\mathrm{CDM}$ cosmology defined by the Planck 2015 results, which defines an analytical mapping between each event's inferred
luminosity distance and it's redshift. While we choose to use MSpline distibutions for the masses and spins, we use the canonical \textsc{PowrerlawRedshift}
model describing the evolution of CBC merger rates as a power law in $(1+z)$ with a slope of $\lambda_z$. The corresponsing distribution is Thus
desribed by $p(z|\lambda)\propto \frac{dV_c}{dz} \frac{T_\mathrm{obs}}{1+z} (1+z)^\lambda$. We model the spin magnitude and tilt distibutions to be 
described by independent MSpline distributions for each component.

>>>>

A common non-parametric method in used in many statistical applications is basis splines. A spline function of order $k$, 
is a piece-wise polynomial of order $k$ polynomials stitched together from defined ``knot'' locations across the domain. 
They provide a useful and cheap way to interpolate generically smooth functions from a finite sampling of ``knot'' heights. 
Basis splines of order $k$ are a set of order $k$ polynomials that form a complete basis for any spline function of order $k$. 
Given an array of knot locations, $\mathbf{t}$ or knot vector, there exists a single unique linear combination of B-Splines for 
every possible spline function with knots, $\mathbf{t}$. One can construct the B-Splines basis components for a given knot vector 
using the Cox-de Boor recursion formula \cite{deBoor78}. Given knots $t_0$, $t_1$,...,$t_{i+k}$ we start with the base case as:

@@

\noindent allowing one to build a basis of $i$ components that spans the vector space of order-$k$ spline functions interpolated 
from this knot vector. This is known as the ``B-Spline'' basis after it's inventor de Boor \cite{deBoor78}. The power of basis splines
comes from the fact that you only have to do the, somewhat expensive, interpolation once for each set of points you want to evaluate your
spline functions at. This allows for much quicker speeds since each evaluation of the spline function becomes a simpler operation that is just
a dot product of 2d matrix and 1d vector. This straightforward operation is also ideal for optimizations from the use GPU accelerators, 
enabling expensive markov chain monte carlo (MCMC) based analyses with hundreds of parameters. Basis splines can easily be generalized to 
their two-dimensional analog, producing tensor product basis splines that, with this computational advantange, allow for high fidelity
modeling of two-dimensional spline functions.

A similar basis, the M-Spline basis \citep{monotone_regression_splines}, has desirable properties when looking to model a probability density function, 
which is the goal of population inference. Namely, $M_i \geq 0$ within $(t_i, t_i+k)$, zero elsewhere, and normalized to unity, $\int M_i(x)dx = 1$. We now have
a useful basis that can represent all spline functions that are probability density functions given our array of knots. The M-Spline basis differs 
from B-Splines by the normalization factor:

@@

\noindent We use this definition of the M-Spline basis for the models presented and applied in the later sections of this paper. 
Figure \ref{fig:spline_basis} shows an example of a 3rd order M-Spline basis with 20 degrees of freedom and knots linearly spaced from 0 to 1.
Another important feature of basis splines is that under appropiate prior conditions, one can alleviate sensitivities to arbitrarily 
chosen prior specifications that splines commonly struggle with. We look to the use of penalized splines or P-Splines \cite{eilers2021practical,BayesianPSplines,Jullion2007RobustSO}, 
that adds a smoothing penalty (or prior) based on the difference of basis spline coeficients. This allows one to much more denseley populate
the domain with knots without the worry of extra variance in the inferred spline functions. We discuss the smoothing prior in more detail 
in Appendix \ref{sec:psplines}, along with our specific prior and basis choices for each model in Appendix \ref{sec:modelpriors}.

### 2022-09-26 src/tex/results.tex
add chi eff plot
methods.tex is mostly finished but w/o citations
Re-work intro, add citations etc

\bruce{intro paragraph here to describe GWTC3 catalog and the selection we made for our analyses}

@@

@@

@@

\begin{figure}
    \script{plot_iid_spintilt.py}
    \includegraphics[width=0.45\textwidth]{figures/iid_spintilt.pdf}
    \caption{Caption.}
    \label{fig:iid_spintilt_dist}
\end{figure}
%\begin{figure*}[ht!]
%    \script{component_spin_distribution_plot.py}
%    \begin{centering}
%%        \includegraphics[width=\textwidth]{figures/component_spin_distribution_plot.pdf}
 %       \caption{Plot showing the inferred component spin magnitude (top) and spin tilt (bottom) distributions inferred with the 9-knot \textsc{MSpline} models. 
%        There are two independent \textsc{MSpline} (black) models for each primary/secondary spin magnitude and spin tilt. We compare by showing the results for the
%        \textsc{Default} (iid) spin model used with the \textsc{PowerlawPeak} mass model (blue) and the \textsc{PowerlawSpline} mass model (orange).}
%        \label{fig:component_spin_distribution}
%    \end{centering}
%\end{figure*}

%\begin{figure*}
%    \script{iid_component_spin_distribution_plot.py}
%    \begin{centering}
%        \includegraphics[width=\textwidth]{figures/iid_component_spin_distribution_plot.pdf}
%        \caption{Plot showing the inferred component spin magnitude (left) and spin tilt (right) distributions inferred with the 9-knot \textsc{MSpline} models. 
%        Each components spin magnitude and tilt are assumed to be IID. We compare by showing the results for the
%        \textsc{Default} (iid) spin model used with the \textsc{PowerlawPeak} mass model (blue) and the \textsc{PowerlawSpline} mass model (orange).}
%        \label{fig:iid_component_spin_distribution}
%    \end{centering}
%\end{figure*}

@@

\subsection{The Effective Spin Dimensions}

@@

>>>>

We use hierarchical Bayesian inference (see Appendix \ref{sec:hierarchical_inference}) to simultaneosly infer the astrophysical mass, spin and redshift distributions of 
a binary black holes (BBHs) given a catalog of gravitational wave observations. We choose to parameterize the binaries' masses with the primary (defined as the more massive component) mass 
and the mass ratio -- defined as $q=\frac{m_2}{m_1}$, bounded in the range from 0 to 1. We choose to model 4 of the 6 total 
spin degrees of freedom of a binary merger with each component spin magnitude $a_1$ and $a_2$, along with the cosine of the tilt angle of each component, 
$\cos{\theta_1}$ and $\cos{\theta_2}$. The tilt angle is defined as the angle between each components' spin vector and the binaries orbital angular mommentum vector. 
We additionally will fit a population model on the redshift or luminosity distance distribution of BBHs. We assume a $\Lambda\mathrm{CDM}$ cosmology using the 
parameters from the Planck 2015 results \cite{Planck2015}, which defines an analytical mapping between each event's inferred
luminosity distance and it's redshift. While we choose to use MSpline distibutions to model the primary mass and spin distributions, we 
model the mass ratio distribution with a power law, i.e. $p(q | m_1, m_\mathrm{min}, \beta_q) \propto q^{\beta_q} \Theta(qm_1 - m_\mathrm{min}) \Theta(m_1 - qm_1)$. 
Where the $\Theta$ is the Heaviside step function which ensures $m_2$ is within the allowed range, [$m_\mathrm{min}$, $m_1$].
We model the evolution of the merger rate with redshift with the \textsc{PowrerlawRedshift} model \cite{Fishbach_2018redshift}, which is desribed by 
$p(z|\lambda_z)\propto \frac{dV_c}{dz} \frac{T_\mathrm{obs}}{1+z} (1+z)^\lambda_z$. We describe our specific prior choices on the hyper-parameters and each model 
in more detail in Appendix \ref{sec:modelpriors}. We choose the same cut on the recent GWTC-3 catalog as was done in the LVK's accompanying population analysis, \bruce{cite}.
Keeping all possible BBH mergers with false alarm rates less than 1 per year, we have 70 oservations. We choose to emit the currently not well understood event, GW190814 \cite{190814disc},
as it was concluded to be an outlier of the rest of the BBH population in both GWTC-2 and GWTC-3 \cite{o3a_pop}. This leaves us with a catlog of 69 BBH mergers observerd oveer a period of about 2 years, from
which we want to infer the population properties of. We start by looking towards the mass distribution.

@@

PLPeak: $m_{1\%} = $\result{$\CIPlusMinus{\macros[MassDistribution][PLPeak][m_1percentile]}$}, $m_{99\%} = $\result{$\CIPlusMinus{\macros[MassDistribution][PLPeak][m_99percentile]}$}

PLSpline: $m_{1\%} = $\result{$\CIPlusMinus{\macros[MassDistribution][PLSpline][m_1percentile]}$}, $m_{99\%} = $\result{$\CIPlusMinus{\macros[MassDistribution][PLSpline][m_99percentile]}$}

MSpline: $m_{1\%} = $\result{$\CIPlusMinus{\macros[MassDistribution][MSpline][m_1percentile]}$}, $m_{99\%} = $\result{$\CIPlusMinus{\macros[MassDistribution][MSpline][m_99percentile]}$}

@@

\subsubsection{Spin Magnitude}

\subsubsection{Spin Orientation}

@@

\begin{figure}
    \script{plot_iid_spintilt.py}
    \includegraphics[width=0.45\textwidth]{figures/iid_spintilt.pdf}
    \caption{Caption.}
    \label{fig:iid_spintilt_dist}
\end{figure}

@@

\subsection{The Effective Spin Dimension}

@@

\begin{figure}
    \script{plot_chieff.py}
    \includegraphics[width=0.45\textwidth]{figures/chi_eff.pdf}
    \caption{Caption.}
    \label{fig:chieff_dist}
\end{figure}
